Cooling Requirements for the Vertical Shear Instability in Protoplanetary Disks

### 2015-05-08 numerical.tex
figure compilation problems (fixed by commenting out for now)
re-shuffled 6.4, updated fig 2, poster files update

\subsection{Vertically non-isothermal disk model} 
We briefly consider a vertically non-isothermal disk model with
$(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma, \Gamma)=(1.4,1.3)$. We set the
vertical domain size close to the physical disk surface, $\zmax =
0.99H_s\simeq 2.6H$, which is smaller than our standard domain
size. In this case the vertical shear rate $\p_z\Omega\propto s$, the
radial entropy gradient. Here, $s=-0.85$.

From the expression for the
critical cooling timescale, Eq. \ref{iso_vsi_cond}, it is reasonable
to expect a similar criteria for vertically non-isothermal 
disks in the form
\begin{align}\label{bcrit_noniso}
 \beta_\mathrm{crit}\to\beta_\mathrm{crit, noniso} =
 \frac{h|s|}{\gamma - \Gamma},
\end{align}
and $\beta_\mathrm{crit,noniso}=0.425$ for the chosen parameters. This
is $\sim 3$ times larger than our fiducial, vertically isothermal
disk with $\beta_\mathrm{crit}=0.125$.

@@

The above calculation demonstrates that the effect of introducing
finite cooling on the VSI operates very similarly in vertically
isothermal disks with an imposed surface, and in vertically
non-isothermal disks with a physical surface. This is despite the fact
that the verticaly buoyancy frequency, $N_z$, formally diverges in
latter case as $|z|\to H_s$.

>>>>

\subsection{Vertically non-isothermal disk model} 
We briefly examine a vertically non-isothermal disk model with 
a physical surface. Recall from Eq. \ref{vertical_shear_ex} that in
general the vertical shear rate $\p_z\Omega\propto s$, the radial
entropy gradient. From the expression for the critical cooling timescale,
Eq. \ref{iso_vsi_cond}, it is reasonable to expect a similar criteria
for vertically non-isothermal disks in the form 
\begin{align}\label{bcrit_noniso}
 \beta_\mathrm{crit}\to\beta_\mathrm{crit, noniso} =
 \frac{h|s|}{\gamma - \Gamma}. 
\end{align}

Here, we consider a disk model with $(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma,
\Gamma)=(1.4,1.3)$. We set the vertical domain size close to the
physical disk surface, $\zmax = 0.99H_s\simeq 2.6H$, which is smaller
than our standard domain size. For this setup, $s=-0.85$ and
$\beta_\mathrm{crit,noniso}=0.425$, which is $\sim 3$ times larger than
our fiducial, vertically isothermal disk with $\beta_\mathrm{crit}=0.125$.

@@

The above calculation demonstrates that the effect of introducing
finite cooling on the VSI operates very similarly in vertically
isothermal disks with an imposed surface, and in vertically
non-isothermal disks with a physical surface. This is despite the fact
that the vertical buoyancy frequency, $N_z$, formally diverges as $|z|\to H_s$.
